### 2014-06-10 ApTranscriptome_TR.Rmd
Analysis with responsive genes separated into 'linear' and 'on' categories

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("RDAVIDWebService")
#biocLite("GOSemSim")
library(topGO) 
library(Rgraphviz)
library(RDAVIDWebService)
library(GOSemSim)

@@

* High - increase expression with temperature
* Low - decrease expression with temperature
* Intermediate - maximum expression at intermediate temperatures (14 - 28C)
* Bimodal - expressed greater than two standard deviations of expression at both low and high temperatures

@@

```{r expression_shape, eval=TRUE, echo=TRUE}

# calculate response type for responsive transcripts
interaction.response.type <- ldply(interaction.lms, .progress = "none", .inform = TRUE, RxNtype)
stopifnot(nrow(interaction.response.type) == length(interaction.lms))

@@

# table
Ap.type.table <- rbind(A22.type.table, Ar.type.table)

@@

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'Ar.type'], Ap.response.type[ , 'A22.type'])
# reorder 
tt2 <- type.table[c("Bimodal", "PositiveLinear", "HighOn", "Intermediate", "NegativeLinear", "LowOn"), c("Bimodal", "PositiveLinear", "HighOn", "Intermediate", "NegativeLinear", "LowOn")]

@@

```{r temperature_response_table, echo=FALSE, results='asis'}
# add row and column totals
tt3 <- rbind(tt2, colSums(tt2))
tt3 <- cbind(tt3, rowSums(tt3))
colnames(tt3)[6] <- "Total"
rownames(tt3)[6] <- "Total"
# move names into table
tt3 <- rbind(rownames(tt3), tt3)
tt3 <- cbind(rownames(tt3), tt3)
# name col/row by species
rownames(tt3) <- c("AcNC", "", "", "", "", "", "")
colnames(tt3) <- c("ApVT", "", "", "", "", "", "")

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), ]

@@

To confirm that there are not sample-level issues, I performed the same comparison using transcripts where I do *not* expect to see a difference in expression.

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in Ar
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), ]
# T test
t.test(log(Ar.int.transcripts$A22.opt+1), log(Ar.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(Ar.int.transcripts$A22.opt, Ar.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)

# list of transcripts that are 'Intermediate' expressed in A22
A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(A22.int.transcripts$A22.opt+1), log(A22.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(A22.int.transcripts$A22.opt, A22.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)

# list of transcripts that are 'Intermediate' expressed in either colony
Ap.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" | Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(Ap.int.transcripts$A22.opt+1), log(Ap.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
w3 <- wilcox.test(Ap.int.transcripts$A22.opt, Ap.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w3
```

@@

```{r extract_Transcripts}
A22.high <- Ap.response.type[which(Ap.response.type$A22.type == "High"), "Transcript"]
Ar.high <- Ap.response.type[which(Ap.response.type$Ar.type == "High"), "Transcript"]

A22.low <- Ap.response.type[which(Ap.response.type$A22.type == "Low"), "Transcript"]
Ar.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), "Transcript"]

A22.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.bim <- Ap.response.type[which(Ap.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), "Transcript"]
Ar.int <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
```

@@

```{r gsea_by_category}
# A22 'High' genes
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high))] <- 1
A22.high.gsea <- gsea(genelist = A22.geneList.high, geneID2GO = geneID2GO)
```

@@

```{r ggplot_high}
trp_A22_high <- ggplot(resp.TPM.dt.sub[A22.high], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp_A22_high)
```

>>>>

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("RDAVIDWebService")
#biocLite("GOSemSim")
library(topGO) 
library(Rgraphviz)
library(GOSemSim)

@@

* HighOn - expression activated at high temperatures
* PositiveLinear - expression positive linear across thermal gradient (0 - 38°C)
* LowOn - expression activated at low temperatures
* NegativeLinear - expression negative linear across thermal gradient (0 - 38°C)
* Intermediate - maximum expression at intermediate temperatures (14 - 28C)
* Bimodal - expressed greater than one standard deviation of expression at both low and high temperatures

@@

```{r expression_shape, eval=TRUE, echo=TRUE}
# calculate response type for responsive transcripts
interaction.response.type <- ldply(interaction.lms, .progress = "none", .inform = TRUE, RxNtype)
stopifnot(nrow(interaction.response.type) == length(interaction.lms))

@@

# table
(Ap.type.table <- rbind(A22.type.table, Ar.type.table))

@@

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'Ar.type'], Ap.response.type[ , 'A22.type'])
# reorder 
tt2 <- type.table[c("Bimodal", "PositiveLinear", "HighOn", "Intermediate", "NegativeLinear", "LowOn", "NotResp"), c("Bimodal", "PositiveLinear", "HighOn", "Intermediate", "NegativeLinear", "LowOn", "NotResp")]

@@

```{r temperature_response_table, echo=FALSE, results='asis'}
# add row and column totals
tt3 <- rbind(tt2, colSums(tt2))
tt3 <- cbind(tt3, rowSums(tt3))
colnames(tt3)[8] <- "Total"
rownames(tt3)[8] <- "Total"
# move names into table
tt3 <- rbind(rownames(tt3), tt3)
tt3 <- cbind(rownames(tt3), tt3)
# name col/row by species
rownames(tt3) <- c("AcNC", "", "", "", "", "", "", "", "")
colnames(tt3) <- c("ApVT", "", "", "", "", "", "", "", "")

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "HighOn" | Ap.response.type$A22.type == "PositiveLinear"), ]

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "LowOn" | Ap.response.type$A22.type == "NegativeLinear"), ]

@@

To confirm that there are not sample-level issues, I performed the same comparison using *Intermediate* expressed-genes where I do not expect to see a difference in expression.

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in either colony
Ap.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" | Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(Ap.int.transcripts$A22.opt+1), log(Ap.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
w3 <- wilcox.test(Ap.int.transcripts$A22.opt, Ap.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w3
```

@@

```{r extract_Transcripts}
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "HighOn"), "Transcript"]
Ar.high.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "HighOn"), "Transcript"]

A22.poslinear.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "PositiveLinear"), "Transcript"]
Ar.poslinear.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "POsitiveLinear"), "Transcript"]

A22.low.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "LowOn"), "Transcript"]
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "LowOn"), "Transcript"]

A22.neglinear.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "NegativeLinear"), "Transcript"]
Ar.neglinear.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "NegativeLinear"), "Transcript"]

A22.bim.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.bim.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), "Transcript"]
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
```

@@

```{r gsea_by_category}
# A22 'High' genes
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high.transcripts))] <- 1
A22.high.gsea <- gsea(genelist = A22.geneList.high, geneID2GO = geneID2GO)
```

@@

```{r ggplot_high}
trp_A22_high <- ggplot(resp.TPM.dt.sub[A22.high.transcripts][1:220,], aes(x=val, y=TPM.scaled, group=Transcript)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp_A22_high)

trp_A22_poslinear <- ggplot(resp.TPM.dt.sub[A22.poslinear.transcripts][1:220,], aes(x=val, y=TPM.scaled, group=Transcript)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp_A22_poslinear)

### predicted values
resp.TPM.dt.sub.pred[,pTPM.scaled:=scale(pTPM), by = Transcript]

pred_A22_high <- ggplot(resp.TPM.dt.sub.pred[A22.high.transcripts][1:220,], aes(x=val, y=pTPM.scaled, group=Transcript)) +
  geom_line() + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(pred_A22_high)

pred_A22_poslinear <- ggplot(resp.TPM.dt.sub.pred[A22.poslinear.transcripts][1:220,], aes(x=val, y=pTPM.scaled, group=Transcript)) +
  geom_line() + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(pred_A22_poslinear)

ggplot(resp.TPM.dt.sub.pred["100348|*|comp3450455_c0_seq1"], aes(x=val, y=pTPM.scaled, group=Transcript)) +
  geom_line() + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))

```
